Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-05-03 Scripts/Manuscript_Draft.Rmd
Modified manuscript for thesis

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "[DO NOT SHARE]"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

* **Corresponding author**: j.palacios@oceans.ubc.ca

@@

We identified 633 transboundary exploited species in the world, more than double the previous estimates [@Teh:2015gd], representing 67.5% of the species analyzed. Between 2005 and 2014, national fleets targeting transboundary species within EEZs cached an average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating USD 65,483 million in fishing revenue. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile, and thus discrete for Ecuador. The contribution of discrete species was 14.8 million tons representing USD 22,167 million in fishing revenue. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary. These species represented only 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level (Fig. 1).

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

img <- readPNG(here("./Figures/Fig1.png"))
grid.raster(img)

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean $\mp$ sd. Revenue in 2010 real USD.](~/GitHub/FishForVisa/Figures/Fig1.png)

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions in terms of revenue (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in E. Asia and the USA (USD 9,605 million) contribute 55% and 82% to the total revenues from the E. Asia and N. America regions, respectively. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. Similarly, Russia shares the Alaskan pollock (*Theragra chalcogramma*) fishery with both Japan and the USA celebrating a management agreement with the latter over the Berring Sea since 1988 [@NOAA:gX6CtUTk]. Transboundary species also contribute largely to fisheries in E. Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] getting one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table S1), however, it has the most revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

```{r Call_Fig_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap=""}

img <- readPNG(here("./Figures/Fig2.png"))
grid.raster(img)
```

![**Figure 2.** Weighted benefits of transboundary species by $kmK2$ and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars $\mp$ s.d.](~/GitHub/FishForVisa/Figures/Fig2.png)
**NOTE to legend: These are not million. They are probably thousand**

We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, category B would be species that are sustainably harvested, and Category C would be species that are over-exploited. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch"}

img <- readPNG(here("./Figures/Fig2.png"))
grid.raster(img)
```

![Figure 3. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](~/GitHub/FishForVisa/Figures/Fig3.png)

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4).

```{r Call_Spp_Circular, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

img <- readPNG(here("./Figures/Fig4.png"))
grid.raster(img)
```

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference’s top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014](~/GitHub/FishForVisa/Figures/Fig4.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. 5). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep].

```{r Call_Fig5, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of shared species by RFMOs. Analysis included with 85 species managed by Fourteen RFMOs"}

img <- readPNG(here("./Figures/Fig5.png"))
grid.raster(img)
```

![Figure 5. Number of shared species across RFMOs. Analysis included 99 species managed by Fourteen RFMOs according to Cullis-Suzuk and Pauly, 2010.](~/GitHub/FishForVisa/Figures/Fig5.png)

On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 3B) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the reconstructed list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their current distributions based on an average from 1970-2000. The total number of species analyzed was 938 representing 54% of the average catch reported by the *Sea Around Us* between 2005 and 2014, including generalized categories such as *Marine fish not identified* or *Mollusca*. However, because we carried the analysis at the species level, we did not include such generalized categories in our comparisons. Our final sample represents 63.2% of identified species, accounting for 96.52% of the catch identified at the species' level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis.

Table 1. Summary of data sources to estimate species' distributions. All data has a resolution  $0.5^{\circ}$ latitude x $0.5^{\circ}$ longitude. All data is publicly available (see Acknowledgments).

```{r Data_Table, eval = T, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  kable_styling(full_width = F) %>%
  column_spec(2,width = "8cm") %>% #wraps text in column2
  column_spec(3,width = "3cm") #%>% #wraps text in column2
  # write_csv("Table_1.csv")

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* consist in the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. This threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Fig. S2).

@@

We estimated the economic contribution in 2010 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was derived by aggregating the catch of species occurring within the EEZs of countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we used the catch data to determine the exploitation category of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies) affect fishing effort and catch volume [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, Category B would be species that are sustainable harvested, and Category C would be species that are over-exploited. Finally, we reported the predominant category over the last ten years of catch data (2005-2014).

@@

Table 2. Rules to determine the category of each analyzed species.

```{r Status_Table, eval = T, echo = F}

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = F) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")

@@

Finally, the results of this study only considered those species that had a *Species Index* of 1 (e.g. all datasets agree on presence) and at least 25% of *Area Index* (e.g. the species shared distribution was at least 25% in each EEZ). This presents a relative conservative estimate of the number of transboundary species in the world. However, using a different *Species Index* will give results based on less datasets while a smaller *Area Index* will estimate more transboundary species (Fig. S2).

<!-- Commented out for Thesis -->
<!-- ## Data Availability -->
<!-- The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa -->

<!-- \clearpage -->

<!-- ## References -->

<!-- <div id="refs"></div> -->

<!-- \clearpage -->

<!-- ## Acknowledgments -->
<!-- We thank T. Clark and A. Tidd for helpful comments and discussions. G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership. -->

<!-- ## Authors contributions -->
<!-- J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis. -->

<!-- ## Additional Information   -->
<!-- **Competing interests**: No declared.  -->

@@

```{r Call_Sup_1, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tons) is represented in the land polygon."}

img <- readPNG(here("./Figures/S1_Catch_EEZ.png"))
grid.raster(img)
```

![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](~/GitHub/FishForVisa/Figures/FigS1.png)

>>>>

---
output:
  bookdown::pdf_book:
    toc : no
    keep_tex: no
    number_sections: yes
    fig_width: 6 
    fig_height: 4 
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
---

@@

@@

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the reconstructed list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their current distributions based on an average from 1970-2000. The total number of species analyzed was 938 representing 54% of the average catch reported by the *Sea Around Us* between 2005 and 2014, including generalized categories such as *Marine fish not identified* or *Mollusca*. However, because we carried the analysis at the species level, we did not include such generalized categories in our comparisons. Our final sample represents 63.2% of identified species, accounting for 96.52% of the catch identified at the species' level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table @\ref(tab:CallTable1)). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis.

<!-- Table 1. Summary of data sources to estimate species' distributions. All data has a resolution  $0.5^{\circ}$ latitude x $0.5^{\circ}$ longitude. All data is publicly available (see Acknowledgments). -->

```{r CallTable1, eval = T, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  kable_styling(full_width = T) %>%
  column_spec(2,width = "2cm") %>% #wraps text in column2
  column_spec(3,width = "5cm") #%>% #wraps text in column2
  # write_csv("Table_1.csv")

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* consist in the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. This threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Figure \@(fig:Call-Sup-2)).

@@

We estimated the economic contribution in 2010 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was derived by aggregating the catch of species occurring within the EEZs of countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we used the catch data to determine the exploitation category of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies) affect fishing effort and catch volume [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table Table \@ref(tab:CallTab2)) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, Category B would be species that are sustainable harvested, and Category C would be species that are over-exploited. Finally, we reported the predominant category over the last ten years of catch data (2005-2014).

@@

<!-- Table 2. Rules to determine the category of each analyzed species. -->

```{r CallTab2, eval = T, echo = F}

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", 
        booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = T) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")

@@

Finally, the results of this study only considered those species that had a *Species Index* of 1 (e.g. all datasets agree on presence) and at least 25% of *Area Index* (e.g. the species shared distribution was at least 25% in each EEZ). This presents a relative conservative estimate of the number of transboundary species in the world. However, using a different *Species Index* will give results based on less datasets while a smaller *Area Index* will estimate more transboundary species (Figure \@(fig:CallSup2)).

We identified 633 transboundary exploited species in the world, more than double the previous estimates [@Teh:2015gd], representing 67.5% of the species analyzed. Between 2005 and 2014, national fleets targeting transboundary species within EEZs cached an average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating USD 65,483 million in fishing revenue. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile, and thus discrete for Ecuador. The contribution of discrete species was 14.8 million tons representing USD 22,167 million in fishing revenue. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary. These species represented only 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level (Figure \@ref(fig:Figure1)).

\clearpage

```{r Figure1, eval = T, fig.width=9, fig.height=12, echo=FALSE, fig.cap="Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

img <- readPNG("~/GitHub/FishForVisa/Figures/Fig1.png")
grid.raster(img)

<!-- ![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean $\mp$ sd. Revenue in 2010 real USD.](~/GitHub/FishForVisa/Figures/Fig1.png) -->

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions in terms of revenue (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Figure \@ref(fig:Figure1)B). China (USD 6,279 million) in E. Asia and the USA (USD 9,605 million) contribute 55% and 82% to the total revenues from the E. Asia and N. America regions, respectively. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Figure \@(fig:CallSup1)). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. Similarly, Russia shares the Alaskan pollock (*Theragra chalcogramma*) fishery with both Japan and the USA celebrating a management agreement with the latter over the Berring Sea since 1988 [@NOAA:gX6CtUTk]. Transboundary species also contribute largely to fisheries in E. Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] getting one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Figure \@ref(fig:CallFigure2)). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table \@ref(tab:TabSup2)), however, it has the most revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

```{r CallFigure2, eval = t, echo=FALSE, fig.pos = "H", fig.cap="Weighted benefits of transboundary species by $kmK2$ and UN sub-region. A) Million 2010 USD. B) Tons. Point sub-region average bars +- s.d."}

img <- readPNG("~/GitHub/FishForVisa/Figures/Fig2.png")
grid.raster(img)
```

<!-- ![**Figure 2.** Weighted benefits of transboundary species by $kmK2$ and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars $\mp$ s.d.](~/GitHub/FishForVisa/Figures/Fig2.png) -->

We determined the exploitation category of each species according to the catch trend within each EEZ (Figure \@ref(fig:CallFigure3)). Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, category B would be species that are sustainably harvested, and Category C would be species that are over-exploited. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table \@ref(tab:TabSup2)). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

```{r CallFigure3, eval = T, fig.width = 10, fig.height = 12, echo=FALSE, fig.cap="Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch"}

img <- readPNG("~/GitHub/FishForVisa/Figures/Fig3.png")
grid.raster(img)
```

<!-- ![Figure 3. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](~/GitHub/FishForVisa/Figures/Fig3.png) -->

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Figure \@ref(fig:CallFigure4)).

```{r CallFigure4, eval = T, fig.width = 10, fig.height = 10, echo=FALSE, fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

img <- readPNG("~/GitHub/FishForVisa/Figures/Fig4.png")
grid.raster(img)
```

<!-- ![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference’s top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014](~/GitHub/FishForVisa/Figures/Fig4.png) -->

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Figure \@ref(fig:CallFigure5)). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep].

```{r CallFigure5, eval = T, echo = FALSE, fig.cap="Number of shared species by RFMOs. Analysis included with 85 species managed by Fourteen RFMOs"}

img <- readPNG("~/GitHub/FishForVisa/Figures/Fig5.png")
grid.raster(img)
```

<!-- ![Figure 5. Number of shared species across RFMOs. Analysis included 99 species managed by Fourteen RFMOs according to Cullis-Suzuk and Pauly, 2010.](~/GitHub/FishForVisa/Figures/Fig5.png) -->

On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Figure \@ref(fig:CallFigure3)B) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

```{r CallSup1, eval = T, fig.width=6, fig.height=4, echo=FALSE, fig.cap="Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tons) is represented in the land polygon."}

img <- readPNG("../Figures/FigS1.png")
grid.raster(img)
```

<!-- ![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](~/GitHub/FishForVisa/Figures/FigS1.png) -->

The number of gridcells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Figure \@(fig:Call-Sup-2)). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Figure \@(fig:CallSup2) darker histogram).

```{r CallSup2, eval = T, fig.width=6, fig.height=4, echo=FALSE, fig.cap="Histogram of transboundary species using different values of Area Index. Darker histogram represents used value (25%) in the analysis"}

img <- readPNG("../Figures/FigS2.png")
grid.raster(img)
```
